---
title: "Robust PCA for functional snippets"
author: "Hyunsung Kim"
date: '2021-04-28'
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    fig_caption: true
    # number_sections: true
    toc: true
    # toc_depth: 2
# header-includes:
#   - \newcommand{\argmin}{\mathop{\mathrm{argmin}}\limits}
---

<style>
  p.caption {   <!-- figure caption -->
    font-size: 0.9em;
    font-style: italic;
    color: grey;
    <!-- margin-right: 10%; -->
    <!-- margin-left: 10%;   -->
    text-align: justify;
  }
  caption {    <!-- table caption -->
    font-size: 0.9em;
    font-style: italic;
    color: grey;
    <!-- margin-right: 10%; -->
    <!-- margin-left: 10%;   -->
    text-align: justify;
  }
</style>


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE, 
  # cache = T,
  fig.align = "center", fig.width = 12, fig.height = 6
)
# Set working directory
# knitr::opts_knit$set(root.dir = "../")
```


# Clustering AMI data

- 3일 평균 데이터로 변환한 2013년 데이터 사용.
- Yao et al.(2005)와 Huber 비교.
- 5개 FPC를 사용하여 $k$-means clustering.
- normalization 과정에서 NA와 0만 관측된 11 가구는 제외.
- $X_i(t) \le 0$ cases (잘못 측정된 경우)
  1. 0는 그냥 두고 음수값 포함된 curve 제거 (**Raw**)
  2. $X_i(t) \le 0$ => NA (**0 - to NA**)
- Normalization
  1. min-max scaling (**min-max**)
  2. divided by a maximum value of each trajectory (**$\max X_i$**)


<br>

# Scatter of 1st FPC vs 2nd FPC

## Raw

```{r scatter, fig.cap = "Figure 1. Scatter of 5 clusters of AMI data. Top is obtained from min-max scaling, and bottom from divided by maximum value of each trajectory.", fig.show = "hold", out.width = "80%"}
path <- "../figure/2021_0428/"
include_graphics(
  paste0(path,
         c("minmax_scatter.png",
           "max(X_i)_scatter.png"))
)
```

## 0 - to NA

```{r scatter_0_to_NA, fig.cap = "Figure 2. Scatter of 5 clusters of AMI data. Top is obtained from min-max scaling, and bottom from divided by maximum value of each trajectory.", fig.show = "hold", out.width = "80%"}
path <- "../figure/2021_0428/"
include_graphics(
  paste0(path,
         c("0_to_NA_minmax_scatter.png",
           "0_to_NA_max(X_i)_scatter.png"))
)
```


<br>


# Average AMI trajectories from total, complete only and missing only

## Raw
```{r each, fig.cap = "Figure 3. Averaged AMI trajectories obtaind from total curves (Black), complete curves (Red) and curves including missing (Blue). Parenthesis indicates the number of curves on each cluster. Top 2 are from min-max scaling, and bottom from divided by a maximum value of each trajectory.", fig.width = 20, fig.height = 14}
load("../RData/20210428_AMI_FPC.RData")
n_group <- 5   # number of clusters
library(tidyverse)

mse_summary <- matrix(0, 2, 4)

# mean trajectories of non-missing and missing for each cluster
par(mfrow = c(4, n_group))
### minmax
fpc.yao <- pca.obj$`minmax`$fpc.yao
fpc.huber <- pca.obj$`minmax`$fpc.huber
AMI_df <- pca.obj$`minmax`$AMI_df
y_outlier <- pca.obj$`minmax`$y_outlier
gr <- seq(0, 1, length.out = ncol(AMI_df))

n_group <- 5   # number of clusters

set.seed(1000)
kmeans.yao.obj <- kmeans(x = fpc.yao, centers = n_group)  
kmeans.yao <- kmeans.yao.obj$cluster

set.seed(1000)
kmeans.huber.obj <- kmeans(x = fpc.huber, centers = n_group)  
kmeans.huber <- kmeans.huber.obj$cluster

# number of curves having missing for each clusters
df <- data.frame(Yao = character(n_group),
                 Huber = character(n_group))
num_missing <- 0
for (i in 1:n_group) {
  ind <- which(kmeans.yao == i)
  y <- AMI_df[ind, ]
  num_NA <- ind[apply(y, 1, function(row) { sum(is.na(row)) }) > num_missing]
  df[i, 1] <- paste(length(num_NA), "/", length(ind))

  ind <- which(kmeans.huber == i)
  y <- AMI_df[ind, ]
  num_NA <- ind[apply(y, 1, function(row) { sum(is.na(row)) }) > num_missing]
  df[i, 2] <- paste(length(num_NA), "/", length(ind))
}

mse_yao <- numeric(n_group)
mse_huber <- numeric(n_group)
for (i in 1:n_group) {
  # cl <- c(2,3,4,1)
  cl <- order(table(kmeans.yao), decreasing = T)
  title <- ifelse(i == 1, 
                  paste0("Yao - Cluster ", i, " (", df[cl[i], 1], ")"),
                  paste0("Cluster ", i, " (", df[cl[i], 1], ")"))
  ind_yao <- which(kmeans.yao == cl[i])
  NA_ind <- which(apply(AMI_df[ind_yao, ], 1, function(row){ sum(is.na(row)) > 0 }))
  matplot(gr,
          cbind(colMeans(AMI_df[ind_yao, ], na.rm = T),
                colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T),
                colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T)),
          type = "l", lwd = c(2, 2, 2), lty = c(1,2,2), col = c(1, 2, 4),
          xlab = "", ylab = "", main = title)
  if (i == 5) {
    legend("topright", 
           c("Total","Complete","Missing"),
           lty = c(1, 1, 1), col = c(2, 2, 4))
  }
  
  # mse
  mse_yao[i] <- cbind(colMeans(AMI_df[ind_yao, ], na.rm = T),
                      colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T),
                      colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T)) %>% 
    apply(2, function(col){
      NA_ind <- which(apply(AMI_df[ind_yao, ], 1, function(row){ sum(is.na(row)) > 0 }))
      c(
        mean((col - colMeans(AMI_df[ind_yao, ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T))^2)
      )
    }) %>% 
    sum() / 2
}
for (i in 1:n_group) {
  # cl <- c(4,3,1,2)
  cl <- order(table(kmeans.huber), decreasing = T)
  title <- ifelse(i == 1, 
                  paste0("Huber - Cluster ", i, " (", df[cl[i], 2], ")"),
                  paste0("Cluster ", i, " (", df[cl[i], 2], ")"))
  ind_huber <- which(kmeans.huber == cl[i])
  NA_ind <- which(apply(AMI_df[ind_huber, ], 1, function(row){ sum(is.na(row)) > 0 }))
  matplot(gr,
          cbind(colMeans(AMI_df[ind_huber, ], na.rm = T),
                colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T),
                colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T)),
          type = "l", lwd = c(2, 2, 2), lty = c(1,2,2), col = c(1, 2, 4),
          xlab = "", ylab = "", main = title)
  
  # mse
  mse_huber[i] <- cbind(colMeans(AMI_df[ind_huber, ], na.rm = T),
                        colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T),
                        colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T)) %>% 
    apply(2, function(col){
      NA_ind <- which(apply(AMI_df[ind_huber, ], 1, function(row){ sum(is.na(row)) > 0 }))
      c(
        mean((col - colMeans(AMI_df[ind_huber, ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T))^2)
      )
    }) %>% 
    sum() / 2
}
mse_summary[1, 1:2] <- c(mean(mse_yao),
                         mean(mse_huber))

### max_Xi
fpc.yao <- pca.obj$`max_Xi`$fpc.yao
fpc.huber <- pca.obj$`max_Xi`$fpc.huber
AMI_df <- pca.obj$`max_Xi`$AMI_df
y_outlier <- pca.obj$`max_Xi`$y_outlier
gr <- seq(0, 1, length.out = ncol(AMI_df))

# Clustering
set.seed(1000)
kmeans.yao.obj <- kmeans(x = fpc.yao, centers = n_group)  
kmeans.yao <- kmeans.yao.obj$cluster

set.seed(1000)
kmeans.huber.obj <- kmeans(x = fpc.huber, centers = n_group)  
kmeans.huber <- kmeans.huber.obj$cluster

# number of curves having missing for each clusters
df <- data.frame(Yao = character(n_group),
                 Huber = character(n_group))
num_missing <- 0
for (i in 1:n_group) {
  ind <- which(kmeans.yao == i)
  y <- AMI_df[ind, ]
  num_NA <- ind[apply(y, 1, function(row) { sum(is.na(row)) }) > num_missing]
  df[i, 1] <- paste(length(num_NA), "/", length(ind))

  ind <- which(kmeans.huber == i)
  y <- AMI_df[ind, ]
  num_NA <- ind[apply(y, 1, function(row) { sum(is.na(row)) }) > num_missing]
  df[i, 2] <- paste(length(num_NA), "/", length(ind))
}

mse_yao <- numeric(n_group)
mse_huber <- numeric(n_group)
for (i in 1:n_group) {
  # cl <- c(2,3,4,1)
  cl <- order(table(kmeans.yao), decreasing = T)
  title <- ifelse(i == 1, 
                  paste0("Yao - Cluster ", i, " (", df[cl[i], 1], ")"),
                  paste0("Cluster ", i, " (", df[cl[i], 1], ")"))
  ind_yao <- which(kmeans.yao == cl[i])
  NA_ind <- which(apply(AMI_df[ind_yao, ], 1, function(row){ sum(is.na(row)) > 0 }))
  matplot(gr,
          cbind(colMeans(AMI_df[ind_yao, ], na.rm = T),
                colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T),
                colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T)),
          type = "l", lwd = c(2, 2, 2), lty = c(1,2,2), col = c(1, 2, 4),
          xlab = "", ylab = "", main = title)
  if (i == 5) {
    legend("topright", 
           c("Total","Complete","Missing"),
           lty = c(1, 1, 1), col = c(2, 2, 4))
  }
  
  # mse
  mse_yao[i] <- cbind(colMeans(AMI_df[ind_yao, ], na.rm = T),
                      colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T),
                      colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T)) %>% 
    apply(2, function(col){
      NA_ind <- which(apply(AMI_df[ind_yao, ], 1, function(row){ sum(is.na(row)) > 0 }))
      c(
        mean((col - colMeans(AMI_df[ind_yao, ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T))^2)
      )
    }) %>% 
    sum() / 2
}
for (i in 1:n_group) {
  # cl <- c(4,3,1,2)
  cl <- order(table(kmeans.huber), decreasing = T)
  title <- ifelse(i == 1, 
                  paste0("Huber - Cluster ", i, " (", df[cl[i], 2], ")"),
                  paste0("Cluster ", i, " (", df[cl[i], 2], ")"))
  ind_huber <- which(kmeans.huber == cl[i])
  NA_ind <- which(apply(AMI_df[ind_huber, ], 1, function(row){ sum(is.na(row)) > 0 }))
  matplot(gr,
          cbind(colMeans(AMI_df[ind_huber, ], na.rm = T),
                colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T),
                colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T)),
          type = "l", lwd = c(2, 2, 2), lty = c(1,2,2), col = c(1, 2, 4),
          xlab = "", ylab = "", main = title)
  
  # mse
  mse_huber[i] <- cbind(colMeans(AMI_df[ind_huber, ], na.rm = T),
                        colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T),
                        colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T)) %>% 
    apply(2, function(col){
      NA_ind <- which(apply(AMI_df[ind_huber, ], 1, function(row){ sum(is.na(row)) > 0 }))
      c(
        mean((col - colMeans(AMI_df[ind_huber, ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T))^2)
      )
    }) %>% 
    sum() / 2
}
mse_summary[1, 3:4] <- c(mean(mse_yao),
                         mean(mse_huber))
par(mfrow = c(1, 1))
```

## 0 - to NA
```{r each_0_to_NA, fig.cap = "Figure 4. Averaged AMI trajectories obtaind from total curves (Black), complete curves (Red) and curves including missing (Blue). Parenthesis indicates the number of curves on each cluster. Top 2 are from min-max scaling, and bottom from divided by a maximum value of each trajectory.", fig.width = 20, fig.height = 14}
n_group <- 5   # number of clusters

# mean trajectories of non-missing and missing for each cluster
par(mfrow = c(4, n_group))
### 0_to_NA_minmax
fpc.yao <- pca.obj$`0_to_NA_minmax`$fpc.yao
fpc.huber <- pca.obj$`0_to_NA_minmax`$fpc.huber
AMI_df <- pca.obj$`0_to_NA_minmax`$AMI_df
y_outlier <- pca.obj$`0_to_NA_minmax`$y_outlier
gr <- seq(0, 1, length.out = ncol(AMI_df))

n_group <- 5   # number of clusters

set.seed(1000)
kmeans.yao.obj <- kmeans(x = fpc.yao, centers = n_group)  
kmeans.yao <- kmeans.yao.obj$cluster

set.seed(1000)
kmeans.huber.obj <- kmeans(x = fpc.huber, centers = n_group)  
kmeans.huber <- kmeans.huber.obj$cluster

# number of curves having missing for each clusters
df <- data.frame(Yao = character(n_group),
                 Huber = character(n_group))
num_missing <- 0
for (i in 1:n_group) {
  ind <- which(kmeans.yao == i)
  y <- AMI_df[ind, ]
  num_NA <- ind[apply(y, 1, function(row) { sum(is.na(row)) }) > num_missing]
  df[i, 1] <- paste(length(num_NA), "/", length(ind))

  ind <- which(kmeans.huber == i)
  y <- AMI_df[ind, ]
  num_NA <- ind[apply(y, 1, function(row) { sum(is.na(row)) }) > num_missing]
  df[i, 2] <- paste(length(num_NA), "/", length(ind))
}

mse_yao <- numeric(n_group)
mse_huber <- numeric(n_group)
for (i in 1:n_group) {
  # cl <- c(2,3,4,1)
  cl <- order(table(kmeans.yao), decreasing = T)
  title <- ifelse(i == 1, 
                  paste0("Yao - Cluster ", i, " (", df[cl[i], 1], ")"),
                  paste0("Cluster ", i, " (", df[cl[i], 1], ")"))
  ind_yao <- which(kmeans.yao == cl[i])
  NA_ind <- which(apply(AMI_df[ind_yao, ], 1, function(row){ sum(is.na(row)) > 0 }))
  matplot(gr,
          cbind(colMeans(AMI_df[ind_yao, ], na.rm = T),
                colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T),
                colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T)),
          type = "l", lwd = c(2, 2, 2), lty = c(1,2,2), col = c(1, 2, 4),
          xlab = "", ylab = "", main = title)
  if (i == 5) {
    legend("topright", 
           c("Total","Complete","Missing"),
           lty = c(1, 1, 1), col = c(2, 2, 4))
  }
  
  # mse
  mse_yao[i] <- cbind(colMeans(AMI_df[ind_yao, ], na.rm = T),
                      colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T),
                      colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T)) %>% 
    apply(2, function(col){
      NA_ind <- which(apply(AMI_df[ind_yao, ], 1, function(row){ sum(is.na(row)) > 0 }))
      c(
        mean((col - colMeans(AMI_df[ind_yao, ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T))^2)
      )
    }) %>% 
    sum() / 2
}
for (i in 1:n_group) {
  # cl <- c(4,3,1,2)
  cl <- order(table(kmeans.huber), decreasing = T)
  title <- ifelse(i == 1, 
                  paste0("Huber - Cluster ", i, " (", df[cl[i], 2], ")"),
                  paste0("Cluster ", i, " (", df[cl[i], 2], ")"))
  ind_huber <- which(kmeans.huber == cl[i])
  NA_ind <- which(apply(AMI_df[ind_huber, ], 1, function(row){ sum(is.na(row)) > 0 }))
  matplot(gr,
          cbind(colMeans(AMI_df[ind_huber, ], na.rm = T),
                colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T),
                colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T)),
          type = "l", lwd = c(2, 2, 2), lty = c(1,2,2), col = c(1, 2, 4),
          xlab = "", ylab = "", main = title)
  
  # mse
  mse_huber[i] <- cbind(colMeans(AMI_df[ind_huber, ], na.rm = T),
                        colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T),
                        colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T)) %>% 
    apply(2, function(col){
      NA_ind <- which(apply(AMI_df[ind_huber, ], 1, function(row){ sum(is.na(row)) > 0 }))
      c(
        mean((col - colMeans(AMI_df[ind_huber, ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T))^2)
      )
    }) %>% 
    sum() / 2
}
mse_summary[2, 1:2] <- c(mean(mse_yao),
                         mean(mse_huber))


### 0_to_NA_max_Xi
fpc.yao <- pca.obj$`0_to_NA_max_Xi`$fpc.yao
fpc.huber <- pca.obj$`0_to_NA_max_Xi`$fpc.huber
AMI_df <- pca.obj$`0_to_NA_max_Xi`$AMI_df
y_outlier <- pca.obj$`0_to_NA_max_Xi`$y_outlier
gr <- seq(0, 1, length.out = ncol(AMI_df))

# Clustering
set.seed(1000)
kmeans.yao.obj <- kmeans(x = fpc.yao, centers = n_group)  
kmeans.yao <- kmeans.yao.obj$cluster

set.seed(1000)
kmeans.huber.obj <- kmeans(x = fpc.huber, centers = n_group)  
kmeans.huber <- kmeans.huber.obj$cluster

# number of curves having missing for each clusters
df <- data.frame(Yao = character(n_group),
                 Huber = character(n_group))
num_missing <- 0
for (i in 1:n_group) {
  ind <- which(kmeans.yao == i)
  y <- AMI_df[ind, ]
  num_NA <- ind[apply(y, 1, function(row) { sum(is.na(row)) }) > num_missing]
  df[i, 1] <- paste(length(num_NA), "/", length(ind))

  ind <- which(kmeans.huber == i)
  y <- AMI_df[ind, ]
  num_NA <- ind[apply(y, 1, function(row) { sum(is.na(row)) }) > num_missing]
  df[i, 2] <- paste(length(num_NA), "/", length(ind))
}

mse_yao <- numeric(n_group)
mse_huber <- numeric(n_group)
for (i in 1:n_group) {
  # cl <- c(2,3,4,1)
  cl <- order(table(kmeans.yao), decreasing = T)
  title <- ifelse(i == 1, 
                  paste0("Yao - Cluster ", i, " (", df[cl[i], 1], ")"),
                  paste0("Cluster ", i, " (", df[cl[i], 1], ")"))
  ind_yao <- which(kmeans.yao == cl[i])
  NA_ind <- which(apply(AMI_df[ind_yao, ], 1, function(row){ sum(is.na(row)) > 0 }))
  matplot(gr,
          cbind(colMeans(AMI_df[ind_yao, ], na.rm = T),
                colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T),
                colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T)),
          type = "l", lwd = c(2, 2, 2), lty = c(1,2,2), col = c(1, 2, 4),
          xlab = "", ylab = "", main = title)
  if (i == 5) {
    legend("topright", 
           c("Total","Complete","Missing"),
           lty = c(1, 1, 1), col = c(2, 2, 4))
  }
  
  # mse
  mse_yao[i] <- cbind(colMeans(AMI_df[ind_yao, ], na.rm = T),
                      colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T),
                      colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T)) %>% 
    apply(2, function(col){
      NA_ind <- which(apply(AMI_df[ind_yao, ], 1, function(row){ sum(is.na(row)) > 0 }))
      c(
        mean((col - colMeans(AMI_df[ind_yao, ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_yao[-NA_ind], ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_yao[NA_ind], ], na.rm = T))^2)
      )
    }) %>% 
    sum() / 2
}
for (i in 1:n_group) {
  # cl <- c(4,3,1,2)
  cl <- order(table(kmeans.huber), decreasing = T)
  title <- ifelse(i == 1, 
                  paste0("Huber - Cluster ", i, " (", df[cl[i], 2], ")"),
                  paste0("Cluster ", i, " (", df[cl[i], 2], ")"))
  ind_huber <- which(kmeans.huber == cl[i])
  NA_ind <- which(apply(AMI_df[ind_huber, ], 1, function(row){ sum(is.na(row)) > 0 }))
  matplot(gr,
          cbind(colMeans(AMI_df[ind_huber, ], na.rm = T),
                colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T),
                colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T)),
          type = "l", lwd = c(2, 2, 2), lty = c(1,2,2), col = c(1, 2, 4),
          xlab = "", ylab = "", main = title)
  
  # mse
  mse_huber[i] <- cbind(colMeans(AMI_df[ind_huber, ], na.rm = T),
                        colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T),
                        colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T)) %>% 
    apply(2, function(col){
      NA_ind <- which(apply(AMI_df[ind_huber, ], 1, function(row){ sum(is.na(row)) > 0 }))
      c(
        mean((col - colMeans(AMI_df[ind_huber, ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_huber[-NA_ind], ], na.rm = T))^2),
        mean((col - colMeans(AMI_df[ind_huber[NA_ind], ], na.rm = T))^2)
      )
    }) %>% 
    sum() / 2
}
mse_summary[2, 3:4] <- c(mean(mse_yao),
                         mean(mse_huber))
par(mfrow = c(1, 1))
```


## MSE between average AMI trajectories in Figure 3 and 4
```{r tab1}
library(kableExtra)
library(tidyverse)

df <- round(mse_summary * 1000, 3)
df <- cbind(c("min-max","$\\max(X_i)$"),
            df)

knitr::kable(df,
             # digits = 3,
             col.names = c("Normalization","Yao","Huber","Yao","Huber"),
             align = "c",
             escape = FALSE,
             caption = "Table 1. MSE between average AMI trajectories in Figure 3 and 4.") %>%
    kable_styling("striped", full_width = FALSE, font_size = 14) %>%
    add_header_above(c(" " = 1,
                       "Raw" = 2,
                       "0 - to NA" = 2))
```


<br>

# Average AMI trajectories for 2 ~ 5 clusters

## Raw

```{r mean, fig.cap = "Figure 5. Averaged AMI trajectories for 2 ~ 5 clusters. Parenthesis indicates the number of curves on each cluster. Top 2 are from min-max scaling, and bottom from divided by a maximum value of each trajectory.", fig.show = "hold", out.width = "100%"}
path <- "../figure/2021_0428/"
include_graphics(
  paste0(path,
         c("minmax_mean.png",
           "max(X_i)_mean.png"))
)
```

## 0 - to NA

```{r mean_0_to_NA, fig.cap = "Figure 6. Averaged AMI trajectories for 2 ~ 5 clusters. Parenthesis indicates the number of curves on each cluster. Top 2 are from min-max scaling, and bottom from divided by a maximum value of each trajectory.", fig.show = "hold", out.width = "100%"}
path <- "../figure/2021_0428/"
include_graphics(
  paste0(path,
         c("0_to_NA_minmax_mean.png",
           "0_to_NA_max(X_i)_mean.png"))
)
```


<br>

# Results summary
- 전체적으로 같은 조건에서 Yao와 Huber의 cluster 결과에 큰 차이가 없음. (아마 outlying 정도가 심하지 않거나 없어서 그런 것으로 보임.)
- PC 2개를 사용한 visualization에서는 $\max X_i$ normalization이 적절해보임. (Figure 1, 2)
- Normalization에 따라 clustering 결과는 달라지지만, 두 방법(Yao, Huber)간 차이는 거의 없음. (Figure 5, 6)


