---
title: "Robust covariance estimation for partially observed functional data"
# author: "Hyunsung Kim"
date: '2021-08-02'
output: 
  pdf_document:
    toc: false
    includes:
      in_header: ./preamble.tex
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE, 
  # cache = T,
  fig.align = "center", 
  out.width = "70%"
  # fig.width = 3, fig.height = 1.5
)
# Set working directory
# knitr::opts_knit$set(root.dir = "../")
```



# 1. Delaigle et al.(2020) setting


```{r, fig.cap="Sample trajectories of Delaigle et al.(2020) setting"}
library(robfpca)
source("../R/sim_delaigle.R")
set.seed(1000)
x.2 <- sim_delaigle(n = 100, 
                    type = "partial",
                    out.prop = 0)
gr <- seq(0, 1, length.out = 51)
x <- list2matrix(x.2)
matplot(gr, t(x), type = "l", xlab = "", ylab = "")
```





\newpage

\begin{table}[ht]
\centering
\footnotesize
\caption{4 FPCs from 50 repetitions}
\begin{tabular}{ccccccccc}
  \hline
   &  & \multicolumn{4}{c}{4 FPCs} & \multicolumn{3}{c}{95\% PVE} \\ 
   \cmidrule(lr){3-6} \cmidrule(lr){7-9}
  Type & Method & PVE & Reconstruction & Completion & Eigenfunction & K & Reconstruction & Completion \\ 
  \hline
  \multirow{11}{*}{Not contaminated } & Yao & 1.00 & \textbf{0.01 (0.01)} & \textbf{0.03 (0.03)} & \textbf{0.06 (0.04)} & 3.94 & \textbf{0.01 (0.02)} & \textbf{0.04 (0.04)} \\ 
  & Huber & 0.84 & 0.08 (0.04) & 0.27 (0.13) & 0.06 (0.09) & 15.40 & 0.11 (0.07) & 0.38 (0.19) \\ 
  & Kraus &  &  & 0.23 (0.21) &  &  &  & 0.23 (0.21) \\ 
  & Kraus-M &  &  & 0.43 (0.25) &  &  &  & 0.43 (0.25) \\ 
  & Kraus-M(sm) &  &  & 1.16 (0.45) &  &  &  & 1.16 (0.45) \\ 
  & Boente & 1.00 & 0.19 (0.21) & 0.56 (0.65) & 0.13 (0.04) & \textbf{3.34} & 0.23 (0.18) & 0.54 (0.50) \\ 
  & M-est & 0.95 & 0.07 (0.04) & 0.23 (0.16) & 0.13 (0.08) & 4.56 & 0.07 (0.04) & 0.25 (0.16) \\ 
  & M-est-noise & 0.95 & 0.04 (0.02) & 0.14 (0.07) & 0.13 (0.08) & 4.52 & 0.04 (0.02) & 0.15 (0.07) \\ 
  & M-est(smooth) & 0.99 & 0.36 (0.74) & 1.07 (2.12) & 0.08 (0.04) & 3.68 & 0.39 (0.74) & 1.11 (2.11) \\ 
  & M-est(smooth)-noise & 0.99 & 0.03 (0.01) & 0.10 (0.04) & 0.08 (0.04) & 3.68 & 0.07 (0.06) & 0.16 (0.10) \\ 
  \hline
  \multirow{11}{*}{Contaminated 1} & Yao & 0.82 & 1.51 (0.25) & 2.10 (0.50) & 1.69 (0.16) & 7.06 & 1.31 (0.28) & 2.01 (0.52) \\ 
  & Huber & 0.84 & 0.09 (0.04) & 0.33 (0.15) & \textbf{0.07 (0.08)} & 15.18 & 0.12 (0.08) & 0.44 (0.21) \\ 
  & Kraus &  &  & 2.64 (0.64) &  &  &  & 2.64 (0.64) \\ 
  & Kraus-M &  &  & 1.35 (0.47) &  &  &  & 1.35 (0.47) \\ 
  & Kraus-M(sm) &  &  & 0.46 (0.29) &  &  &  & 0.46 (0.29) \\ 
  & Boente & 1.00 & 0.11 (0.13) & 0.31 (0.46) & 0.15 (0.05) & 3.78 & 0.13 (0.13) & 0.32 (0.37) \\ 
  & M-est & 0.68 & 0.16 (0.12) & 0.46 (0.39) & 0.21 (0.13) & 18.80 & 0.18 (0.16) & 0.62 (0.59) \\ 
  & M-est-noise & 0.71 & 0.10 (0.04) & 0.26 (0.12) & 0.21 (0.13) & 15.80 & \textbf{0.06 (0.02)} & \textbf{0.23 (0.10)} \\ 
  & M-est(smooth) & 0.99 & 2.58 (3.84) & 8.32 (11.93) & 0.09 (0.05) & 3.80 & 2.56 (3.79) & 8.14 (11.68) \\ 
  & M-est(smooth)-noise & 0.99 & \textbf{0.06 (0.02)} & \textbf{0.19 (0.08)} & 0.09 (0.05) & \textbf{3.60} & 0.11 (0.07) & 0.26 (0.13) \\ 
  \hline
  \multirow{14}{*}{Contaminated 2} & Yao & 0.94 & 0.69 (0.53) & 1.20 (0.70) & 1.26 (0.59) & 4.38 & 0.66 (0.58) & 1.15 (0.73) \\ 
  & Huber & 0.84 & 0.10 (0.04) & 0.38 (0.17) & \textbf{0.07 (0.09)} & 15.40 & 0.14 (0.09) & 0.49 (0.23) \\ 
  & Kraus &  &  & 1.95 (0.89) & &  &  & 1.95 (0.89)  \\ 
  & Kraus-M &  &  & 1.04 (0.42) & &  &  & 1.04 (0.42)  \\ 
  & Kraus-M(sm) &  &  & 0.49 (0.27) & &  &  & 0.49 (0.27)  \\ 
  & Boente & 1.00 & 0.21 (0.55) & 0.60 (1.42) & 0.14 (0.05) & 3.84 & 0.23 (0.54) & 0.60 (1.39) \\ 
  & M-est & 0.76 & 0.20 (0.40) & 0.66 (1.42) & 0.19 (0.13) & 16.90 & 0.26 (0.46) & 0.90 (1.65) \\ 
  & M-est-noise & 0.82 & 0.07 (0.02) & 0.24 (0.10) & 0.19 (0.13) & 11.40 & \textbf{0.06 (0.02)} & \textbf{0.24 (0.10)} \\ 
  & M-est(smooth) & 0.99 & 0.93 (1.50) & 3.06 (5.12) & 0.10 (0.07) & 3.70 & 0.95 (1.46) & 3.01 (4.98) \\ 
  & M-est(smooth)-noise & 0.99 & \textbf{0.06 (0.02)} & \textbf{0.20 (0.09)} & 0.10 (0.07) & \textbf{3.48} & 0.12 (0.07) & 0.29 (0.14) \\ 
  & {\color{blue} Kraus-M(sm)} &  &  & 0.24 (0.13) &  &&& \\ 
  & {\color{blue} M-est(smooth)} & 0.97 & 0.07 (0.05) & 0.24 (0.14) & 0.14 (0.10) &&& \\ 
  & {\color{blue} M-est(smooth)-noise} & 0.97 & 0.06 (0.02) & 0.20 (0.09) & 0.14 (0.10) &&& \\ 
  \hline
\end{tabular}
\end{table}


\newpage

# 2. Kraus(2015) setting

```{r, fig.cap="Sample trajectories of Kraus(2015) setting"}
library(robfpca)
source("../R/sim_kraus.R")
set.seed(1000)
x.2 <- sim_kraus(n = 100, 
                 type = "partial",
                 out.prop = 0)
gr <- seq(0, 1, length.out = 51)
x <- list2matrix(x.2)
matplot(gr, t(x), type = "l", xlab = "", ylab = "")
```



\begin{table}[ht]
\centering
\footnotesize
\caption{5 FPCs from 50 repetitions}
\begin{tabular}{ccccccccc}
  \hline
   &  & \multicolumn{4}{c}{5 FPCs} & \multicolumn{3}{c}{95\% PVE} \\ 
   \cmidrule(lr){3-6} \cmidrule(lr){7-9}
  Type & Method & PVE & Reconstruction & Completion & Eigenfunction & K & Reconstruction & Completion \\ 
  \hline
  \multirow{11}{*}{Not contaminated} & Yao & 1.00 & \textbf{0.01 (0.00)} & \textbf{0.02 (0.01)} & \textbf{0.12 (0.12)} & 3.04 & \textbf{0.02 (0.00)} & \textbf{0.03 (0.01)} \\ 
  & Huber & 0.85 & 0.06 (0.03) & 0.18 (0.08) & 1.42 (0.07) & 15.20 & 0.08 (0.04) & 0.22 (0.09) \\ 
  & Kraus &  &  & \textbf{0.02 (0.01)} & &  &  & 0.02 (0.01)  \\ 
  & Kraus-M &  &  & 0.06 (0.02) & &  &  & 0.06 (0.02)  \\ 
  & Kraus-M(sm) &  &  & 0.09 (0.04) &  &  &  & 0.09 (0.04) \\ 
  & Boente & 1.00 & 0.05 (0.03) & 0.14 (0.08) & 0.74 (0.13) & \textbf{2.94} & 0.07 (0.02) & 0.13 (0.06) \\ 
  & M-est & 0.94 & 0.02 (0.01) & 0.05 (0.02) & 0.34 (0.19) & 6.18 & 0.02 (0.01) & 0.06 (0.02) \\ 
  & M-est-noise & 0.94 & 0.02 (0.00) & 0.04 (0.01) & 0.34 (0.19) & 5.94 & 0.01 (0.00) & 0.04 (0.02) \\ 
  & M-est(smooth) & 0.99 & 0.06 (0.07) & 0.17 (0.19) & 0.45 (0.19) & 3.18 & 0.05 (0.04) & 0.09 (0.07) \\ 
  & M-est(smooth)-noise & 0.99 & 0.02 (0.00) & 0.04 (0.02) & 0.45 (0.19) & 3.16 & 0.03 (0.01) & 0.04 (0.01) \\ 
  \hline
  \multirow{11}{*}{Contaminated 1} & Yao & 0.89 & 0.44 (0.08) & 0.59 (0.14) & 1.64 (0.15) & 6.72 & 0.42 (0.08) & 0.59 (0.14) \\ 
  & Huber & 0.85 & 0.09 (0.04) & 0.25 (0.09) & 1.42 (0.08) & 15.28 & 0.10 (0.04) & 0.29 (0.10) \\ 
  & Kraus &  &  & 1.01 (0.29) & &  &  & 1.01 (0.29)  \\ 
  & Kraus-M &  &  & 0.17 (0.06) &  &  &  & 0.17 (0.06) \\ 
  & Kraus-M(sm) &  &  & 0.06 (0.02) & &  &  & 0.06 (0.02)  \\ 
  & Boente & 1.00 & 0.10 (0.14) & 0.33 (0.40) & 0.69 (0.12) & \textbf{2.92} & 0.11 (0.11) & 0.24 (0.26) \\ 
  & M-est & 0.70 & 0.05 (0.04) & 0.10 (0.09) & 0.72 (0.12) & 20.22 & 0.08 (0.09) & 0.26 (0.31) \\ 
  & M-est-noise & 0.77 & 0.03 (0.01) & \textbf{0.05 (0.01)} & 0.72 (0.12) & 14.18 & \textbf{0.02 (0.00)} & \textbf{0.05 (0.01)} \\ 
  & M-est(smooth) & 0.99 & 0.17 (0.17) & 0.52 (0.58) & 0.49 (0.14) & 3.52 & 0.14 (0.15) & 0.34 (0.46) \\ 
  & M-est(smooth)-noise & 1.00 & \textbf{0.02 (0.00)} & \textbf{0.05 (0.01)} & \textbf{0.44 (0.15)} & 3.06 & 0.03 (0.01) & \textbf{0.05 (0.01)} \\ 
  \hline
  \multirow{14}{*}{Contaminated 2} & Yao & 0.97 & 0.29 (0.18) & 0.35 (0.22) & 1.61 (0.26) & 3.90 & 0.29 (0.18) & 0.36 (0.22) \\ 
  & Huber & 0.86 & 0.10 (0.10) & 0.25 (0.14) & 1.42 (0.09) & 13.94 & 0.12 (0.11) & 0.30 (0.16) \\ 
  & Kraus &  &  & 0.62 (0.66) & &  &  & 0.62 (0.66)  \\ 
  & Kraus-M &  &  & 0.15 (0.05) & &  &  & 0.15 (0.05)  \\ 
  & Kraus-M(sm) &  &  & 0.05 (0.02) & &  &  & 0.05 (0.02)  \\ 
  & Boente & 1.00 & 0.07 (0.07) & 0.22 (0.20) & 0.76 (0.10) & \textbf{2.98} & 0.09 (0.05) & 0.17 (0.13) \\ 
  & M-est & 0.73 & 0.04 (0.03) & 0.08 (0.06) & 0.67 (0.11) & 19.32 & 0.07 (0.08) & 0.22 (0.25) \\ 
  & M-est-noise & 0.80 & 0.03 (0.01) & 0.05 (0.01) & 0.67 (0.11) & 13.54 & \textbf{0.02 (0.01)} & \textbf{0.04 (0.01)} \\ 
  & M-est(smooth) & 0.99 & 0.09 (0.09) & 0.27 (0.28) & 0.46 (0.15) & 3.36 & 0.08 (0.07) & 0.17 (0.20) \\ 
  & M-est(smooth)-noise & 1.00 & \textbf{0.02 (0.01)} & \textbf{0.04 (0.01)} & \textbf{0.41 (0.15)} & 3.06 & 0.03 (0.01) & 0.05 (0.01) \\ 
  & {\color{blue} Kraus-M(sm)} &  &  & 0.04 (0.01) &  &&& \\ 
  & {\color{blue} M-est(smooth)} & 0.98 & 0.02 (0.01) & 0.05 (0.02) & 0.34 (0.12) &&& \\ 
  & {\color{blue} M-est(smooth)-noise} & 0.98 & 0.02 (0.00) & 0.04 (0.01) & 0.34 (0.12) &&& \\ 
  \hline
\end{tabular}
\end{table}



