---
title: "Bootstrap aggregated sparse FPCA for classification"
author: "Hyunsung Kim"
date: January 31, 2020
institute: Department of Statistics \newline Chung-Ang University
fonttheme: "professionalfonts"
output:
  beamer_presentation:
    theme: "metropolis"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("C:\\Users\\user\\Desktop\\KHS\\FDA-Lab\\Sparse FPCA\\Application\\bootstrap_aggregating")
load("result.RData")
library(tidyverse)
library(gridExtra)
```


## Simulation

- The bone mineral density data
  - 280 observations(186 training set and 94 test set)
  - Sparse data
  - Fit the classifiers for 30 different seeds
  

## Simulation results

\begin{table}[ht]
\caption{The mean accuracy for different 30 seeds}
\centering
\tiny
\begin{tabular}{ccccccc}
  \hline
  & Logit & SVM(Gaussian) & SVM(Linear) & SVM(Sigmoid) & SVM(Poly) \\ 
  \hline
  Single & 0.684 & 0.667 & 0.678 & 0.663 & 0.626 \\ 
  Majority & \textcolor{red}{0.690} & \textcolor{red}{0.683} & 0.685 & \textcolor{red}{0.682} & \textcolor{red}{0.660} \\ 
  OOB & 0.688 & 0.681 & \textcolor{red}{0.686} & 0.680 & 0.658 \\ 
  \hline
\end{tabular}
\end{table}


## Simulation results

```{r echo=F, fig.cap="Accuracy of classifiers for different seeds", out.width="300px", out.height="200px"}
cname <- c("Logit", "SVM(Gaussian)", NA, "SVM(Linear)", "SVM(Sigmoid)", "SVM(Poly)", NA)
for (i in c(1,2,4,5,6)) {
  assign(paste("p", i, sep=""),
         result[[i]] %>%
  reshape2::melt(id="seed", value.name="Accuracy") %>%
  ggplot(aes(seed/100, Accuracy, group=variable, colour=variable)) +
  geom_line() +
  ylim(0.5, 0.9) +
  xlab("") +
  ylab("Accuracy") +
  ggtitle(cname[i]) +
  theme_bw() +
  theme(legend.title = element_blank()))
}
grid.arrange(p1, p2, p4, p5, p6, ncol=3)
```












